Laplacian envelope

### 2021-07-31 Notes/laplace.tex
fixed labels in spacing fig

The derivatives, for \(m \neq n\) are
%
\begin{equation}\label{eq:finitederivs}
\begin{aligned}
  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
    = 0.
\end{aligned}
\end{equation}
%
We can solve for the KKT multipliers by noting that
%
\begin{equation*}
  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
  \qquad \text{and} \qquad
  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
\end{equation*}
%
Then, with some new vectors defined:
%
\begin{equation}\label{eq:kktsol}
\begin{aligned}
  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
    + \lambda \ombm^\nu_m(0) 
    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
    }, \\
  \mu^\nu_{mn} &= \eqm_m \big( 
      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
  \\ &\phantom{= \eqm_m \big( }
    + \brk{{c}_m(s) + (-1)^\nu} 
        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
     \big), \\
  &\text{where} \quad 
  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
\end{aligned}
\end{equation}
%
If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2f^\nu \), we find that
%
\begin{equation}\label{eq:diagKKT}
\begin{gathered}
  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}{2f^\nu},
  \qquad 
  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}{2 f^\nu}, \\
\begin{aligned}
  \mu^\nu_{mm} &= - \frac{\eqm_m}{2f^\nu} 
    \left[ 
      c_m(s) (\delta \wm_m + s \delta \etwm_m(s)) - \hat{A}(s)
      + \lambda \delta w_m
      \right. \\ &\phantom{= - \frac{\eqm_m}{2f^\nu} []}\left.
      + (-1)^\nu \prn{ 
        [\encm \thbm(s)]_m - c_m(s) a_m(s)
        + \delta \wm_m + s \delta \etwm_m(s)
        - \lambda a_m(0)
      }
    \right], \\
\end{aligned}
\end{gathered}
\end{equation}
%
where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).

@@

The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
%
\begin{equation}\label{eq:shiftscale}
  \zeta {\circledast}_{s} \MMd 
      \equiv \zeta \MMd
      + (1 - \zeta) \brk{(1 + s)\I - s \onev \eq} .
\end{equation}
%
Then we have
%
\begin{equation}\label{eq:lshiftscale}
  \hat{A} \prn{s; \MMd^\nu} 
    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
\end{equation}
%
The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
is not invariant under this transformation, unlike the \(s = 0\) case.
Using this transformation to fix the upper bound can break the lower bound.

@@

For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
%
\begin{equation}\label{eq:separator}
  \forall\ m, \;\; \exists\ n : 
  \begin{cases}
    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
  \end{cases}
\end{equation}
%
We have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_pert} \aligntop{%
  \includegraphics[width=0.28\linewidth]{pert_triangle.svg}}
  \item \label{fig:triangle_left} \aligntop{%
  \includegraphics[width=0.28\linewidth]{left_triangle.svg}}
  \item \label{fig:triangle_right} \aligntop{%
  \includegraphics[width=0.28\linewidth]{right_triangle.svg}}
  \\[1em]
  \item \label{fig:triangle_allowed} \aligntop{%
  \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
  \item \label{fig:triangle_spacing} \aligntop{%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
\end{myenuma}
\caption[Triangularising perturbation]{Triangularising perturbation.
(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}.
(\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
(\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
\label{fig:triangle_s}}
\end{center}
\end{figure}

@@

Let us define the following
%
\begin{equation}\label{eq:degenerate}
  \dgn_{m} = \thb + c_m \etw,
  \quad
  \dgn^\nu_m = \dgn_m + (-1)^\nu \etw,
  \quad \implies \;
  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
\end{equation}
%
They should be thought of as collections of vectors rather than matrices.

@@

Both \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
Such a symmetry would imply that the states are lumpable.
This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.

@@

The area (and \(\Delta\pr\)) is invariant under the scale transformation
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv \zeta \MMd + (1-\zeta)\I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.

>>>>

The derivatives, for \(m \neq n\) are
%
\begin{equation}\label{eq:finitederivs}
\begin{aligned}
  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = & \,
    \eqm_m \prn{ \brk{\thbm_m(s) - \thbm_n(s)}
     + \prn{\mathbf{c}_m(s) + (-1)^\nu} \brk{\etwm_m(s) - \etwm_n(s)} } \\
     &+ \mu^\nu_{mn} -  \mu^\nu_{mm}
     + \lambda \eqm_m (\etwm_{m}(0)-\etwm_{n}(0)) 
    = 0.
\end{aligned}
\end{equation}
%
We can solve for the KKT multipliers by noting that
%
\begin{equation*}
  \mu^\nu_{ij} \MMdm^\nu_{ij} = 0,
  \qquad \text{and} \qquad
  \sum_{j: j \neq i} \mu^\nu_{ii} \MMdm^\nu_{ij} 
      = \mu^\nu_{ii} (1 - \MMdm^\nu_{ii}) = \mu^\nu_{ii}.
\end{equation*}
%
Then, with some new vectors defined:
%
\begin{equation}\label{eq:kktsol}
\begin{aligned}
  \mu^\nu_{mm} &= - \eqm_m \prn{ \albm^\nu_m(s) 
    + \lambda \ombm^\nu_m(0) 
    + \prn{{c}_m(s) + (-1)^\nu} \ombm^\nu_m(s)
    }, \\
  \mu^\nu_{mn} &= \eqm_m \big( 
      \brk{ \thbm_n(s) - \thbm_m(s) - \albm^\nu_m(s) }
    + \lambda \brk{ \etwm_n(0) - \etwm_m(0) - \ombm^\nu_m(0) } 
  \\ &\phantom{= \eqm_m \big( }
    + \brk{{c}_m(s) + (-1)^\nu} 
        \brk{ \etwm_n(s) - \etwm_m(s) - \ombm^\nu_m(s) }
     \big), \\
  &\text{where} \quad 
  \alb^\nu(s) = (\MMd^\nu - \I) \thb(s), \qquad
  \omb^\nu(s) = (\MMd^\nu - \I) \etw(s).
\end{aligned}
\end{equation}
%
If we note that we can write \( \MMd^\nu - \I = \prn{\frg + (-1)^\nu \enc} / 2f^\nu \), we find that
%
\begin{equation}\label{eq:diagKKT}
\begin{gathered}
  \alb^\nu(s) = \frac{\delta \mathbf{a}(s) + (-1)^\nu \enc \thb(s)}
                     {2f^\nu},
  \qquad 
  \omb^\nu(s) = \frac{\delta \w + s \delta \etw(s) - (-1)^\nu \mathbf{a}(s)}
                     {2 f^\nu}, \\
\begin{aligned}
  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
    \left[ 
      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
      - \lambda \delta w_m
    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
    \left.
      + (-1)^\nu \prn{ 
        c_m(s) a_m(s) - [\encm \thbm(s)]_m
        - \delta \wm_m - s \delta \etwm_m(s)
        + \lambda a_m(0)
      }
    \right], \\
\end{aligned}
\end{gathered}
\end{equation}
%
where \(\delta\) indicates mean subtraction and \(\eq \mathbf{a}(s) = \hat{A}(s)\).

@@

The corresponding symmetry here is scaling the shifted process from \cref{eq:lshiftpd}, which amounts to
%
\begin{equation}\label{eq:shiftscale}
  \zeta {\circledast}_{s} \MMd 
      \equiv (1 - \zeta) \MMd
      + \zeta \brk{(1 + s)\I - s \onev \eq} .
\end{equation}
%
Then we have
%
\begin{equation}\label{eq:lshiftscale}
  \hat{A} \prn{s; \MMd^\nu} 
    = \hat{A} \prn{0, \shift{\MMd}^\nu(s)}
    = \hat{A} \prn{0, \zeta * \shift{\MMd}^\nu(s)}
    = \hat{A} \prn{s, \zeta \circledast_s \MMd^\nu}.
\end{equation}
%
The problem here is that the lower bound, \(\MMdm^\nu_{ij} \geq 0\), 
is not invariant under this transformation, unlike the \(s = 0\) case.
Using this transformation to fix the upper bound can break the lower bound.

@@

For any state \(m\), there is a state \(n\) such that all potentiating transitions go to equal or smaller values of \(\etw\) all depressing transitions go to equal or greater values.
%
\begin{equation}\label{eq:separator}
  \forall\ m, \;\; \exists\ n : 
  \begin{cases}
    \etwm_i > \etwm_n &\implies \MMdm\pot_{mi} = 0, \\
    \etwm_i < \etwm_n &\implies \MMdm\dep_{mi} = 0.
  \end{cases}
\end{equation}
%
The remaining allowed transitions from \(m\) are depicted in \cref{fig:triangle_s}(\ref{fig:triangle_allowed}).
Note that we have not yet ruled the possibility that both \( \MMdm\pot_{mn} \) and \( \MMdm\dep_{mn} \) are nonzero.
If one of those transitions is zero, there will be two states with the property above: the targets of right-most depressing and left-most potentiating transitions.

@@

\begin{figure}[ht]
\begin{center}
\begin{myenuma}
  \item \label{fig:triangle_pert} \aligntop{%
  \includegraphics[width=0.28\linewidth]{pert_triangle.svg}}
  \item \label{fig:triangle_left} \aligntop{%
  \includegraphics[width=0.28\linewidth]{left_triangle.svg}}
  \item \label{fig:triangle_right} \aligntop{%
  \includegraphics[width=0.28\linewidth]{right_triangle.svg}}
  \\[1em]
  \item \label{fig:triangle_allowed} \aligntop{%
  \includegraphics[width=0.4\linewidth]{allowed_triangle.svg}}
  \item \label{fig:triangle_spacing} \aligntop{%
  \includegraphics[width=0.4\linewidth]{spacing_triangle.svg}}
\end{myenuma}
\caption[Triangularising perturbation]{Triangularising perturbation.
(\ref*{fig:triangle_pert}) The type of perturbation considered in \cref{eq:pertKupdn}, indicating which transitions have their rates increased/decreased.
(\ref*{fig:triangle_left},\ref*{fig:triangle_right}) Possible end-points of this perturbation. 
(\ref*{fig:triangle_allowed}) The not-as-yet ruled out transitions when states are sorted so that \(\etw\) decreases to the right, with \(\etwm_m\) unconstrained (as yet).
(\ref*{fig:triangle_spacing}) The consequences of \cref{eq:oneexit,eq:doubleexit,eq:twoexits,eq:twodoubleexits,eq:nonexit} in the \( (\dgnm_{m,l},-\etwm_l) \)-plane, with \(i\) a state with a nonzero depressing transition from \(m\), \(j\) a state with a nonzero potentiating transition, \(n\) a state with both depressing and potentiating transitions and \(k\) a state with no nonzero transitions from \(m\).
The positions of \(m\) and \(k\) are not yet constrained, other than being in the upper quadrant \wrt \(n\).
\label{fig:triangle_s}}
\end{center}
\end{figure}

@@

Let us define the following
%
\begin{equation}\label{eq:degenerate}
  \dgn_{m} = \thb + c_m \etw,
  \quad
  \dgn^\nu_m = \dgn_m + (-1)^\nu \etw,
  \quad \implies \;
  \frac{1}{f^\nu} \pdiff{\CL}{\MMdm^\nu_{mn}} = \eqm_m (\dgnm^\nu_{m,m} - \dgnm^\nu_{m,n}) + \mu^\nu_{mn} - \mu^\nu_{mm}.
\end{equation}
%
They should be thought of as collections of vectors rather than matrices.
We can also write \cref{eq:diagKKT} as
%
\begin{equation}\label{eq:diagKKTz}
\begin{aligned}
  \mu^\nu_{mm} &= \frac{\eqm_m}{2f^\nu} 
    \left[ 
      \hat{A}(s) - c_m(s) (\delta \wm_m + s \delta \etwm_m(s))
      - \lambda \delta w_m
    \right. \\ &\phantom{= \frac{\eqm_m}{2f^\nu} }
    \left.
      - (-1)^\nu \prn{ 
        [\encm \dgnm(s)]_m
        + \delta \wm_m + s \delta \etwm_m(s)
        - \lambda a_m(0)
      }
    \right], \\
\end{aligned}
\end{equation}
%

@@

Both of \cref{eq:twoexits,eq:twodoubleexits} require degeneracy.
The physicist's intuition states that degeneracy is fragile and requires a symmetry to protect it.
Such a symmetry would imply that the states are lumpable.
This would imply that, after lumping all lumpable states together, the situation in \cref{eq:twoexits} is impossible and \cref{eq:twodoubleexits} is even more impossible.
If this intuition holds, we expect that \cref{fig:triangle_s}(\ref{fig:triangle_spacing}) would have at most one state on each of the two diagonal lines.
Either there would be one state like \(n\), or there would be one state like \(i\) and one state like \(j\).

@@

The area (and \(\Delta\pr\)) is invariant under the scale transformation
%
\begin{equation}\label{eq:scale}
  \zeta * \MMd \equiv (1-\zeta) \MMd + \zeta \I .
\end{equation}
%
It follows that
%
\begin{equation*}
  \sum_{\nu, m \neq n} \MMdm^\nu_{mn} \pdiff{\CL}{\MMdm^\nu_{mn}} = - \sum_{\nu, m} \mu^\nu_{mm} = 0 .
\end{equation*}
%
As the KKT multipliers are nonnegative, this implies that all the diagonal KKT multipliers vanish and we can ignore the diagonal constraints.
